
<h3>Nonlinear regression with the linear model</h3>
<p>Start by loading the data</p>

```{r}
data <- read.csv("../../data/olympic100m.txt", header=FALSE)
x <- data$V1
t <- data$V2
```

<h4>Rescaling $x$</h4>
<p>We re-scale the Olympic year value to help with numerical problems when we
look at high orders. Satisfy yourself that this doesn't make any difference!</p>

```{r}
x <- (x - 1896) / 4
```

<h4>Creating $\mathbf{X}$</h4>
<p>We loop up to the desired order, adding columns to $\mathbf{X}$ at each
stage</p>

```{r}
maxorder <- 5
X <- rep(1, length(x))
for (i in 1:maxorder)
    X <- cbind(X, (x**i))
```
    
<p>Solve the system...</p>

```{r}
w <- solve(t(X) %*% X, t(X) %*% t)
print(w)
```

Create a test set in the same manner as $\mathbf{X}$...

```{r}
x_test <- seq(0, 30, length=100)
X_test <- rep(1, length(x_test))
for (i in 1:maxorder)
    X_test <- cbind(X_test, x_test**i)
f_test <- X_test %*% w
```

```{r}
plot(x, t, col="red", pch=16)
lines(x_test, f_test, col="blue")
```

<p>Try playing with the maximum order and extending the test data range...</p>
