```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

<h2>Generative model for the Olympic 100m</h2>
<p>We will now look at generating data that looks a bit like the Olympic 100m
data. We'll start by getting the data.</p>

```{r}
data <- read.csv("../../data/olympic100m.txt", header=FALSE)
x <- data$V1
t <- data$V2
```

<p>We'll assume that we know the true parameter values and look at generating
the noise. We'll get the true parameter values by minimising the loss...</p>

```{r}
X <- cbind(rep(1, length(x)), x)
w <- solve(t(X) %*% X, t(X) %*% t)
print(w)
```


<p>Choose values for the mean and variance of the noise</p>

```{r}
mu <- 0.0
sigma_sq <- 0.05
```

<p>Generate the noisy t values (note that `rnorm` requires the
standard deviation, not the variance)</p>

```{r}
noisy_t <- X %*% w + rnorm(length(t), mean=mu, sd=sqrt(sigma_sq))
```

<p>Plot the noisy data</p>

```{r}
plot(x, noisy_t, col="black", pch=16)
lines(x, X %*% w, col="blue")
```

<p>A second plot, this time with the real data included</p>

```{r}
plot(x, noisy_t, col="black", pch=16)
points(x,t, col = "red", pch=16)
lines(x, X %*% w, col="blue")
```